---
title: "R Notebook"
author: Caroline Sanicola
date: August 26, 2023
output: 
    html_notebook:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: true
    theme: slate
    highlight: default
---

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(ggpubr)
library(ggsci)
library(ggthemes)
library(ggExtra)
library(ggcorrplot)
library(ggfortify)
library(ggrepel)
```

# Introduction to R
# 1. Calculations
```{r}
25 * 4 + 9 / 3
```

# 2. Comments
```{r}
# this is how to comment in an R code
3 * 3 * 3
```

# 3. Data Types
```{r}
class(2) # numeric
class("hello") # character
class("23") # character
class(FALSE) # logical
class(NA) # logical
```

# 4. Variables
We assign a value to a variable using the assignment operator `<-`
```{r}
x <- 3
x

full_name <- "Caroline Sanicola"
print(full_name)

# Greeting
message_string <- "Hello there"
print(message_string)

# Farewell
message_string <- "Hasta la vista"
print(message_string)
```

# 5. Vectors
List-like structure that contain items of the same data type.
```{r}
spring_months <- c("March", "April", "May", "June")
spring_months

# Vector of numbers
numbers <- c(1, 2, 3, 4, 5)
numbers
```

# 6. Conditionals
```{r}
if (TRUE) {
    print("Hello")
}

if (FALSE) {
    print("Hello")
}

# if else
if (TRUE) {
    print("Hello")
} else {
    print("Goodbye")
}

# if else if
if (TRUE) {
    print("Hello")
} else if (FALSE) {
    print("Goodbye")
} else {
    print("See you later")
}
```
```{r}
# if statement
x <- 5
if (x > 3) {
    print("x is greater than 3")
}

# if else statement
x <- 2
if (x > 3) {
    print("x is greater than 3")
} else {
    print("x is less than or equal to 3")
}

# if else if statement
x <- 3
if (x > 3) {
    print("x is greater than 3")
} else if (x < 3) {
    print("x is less than 3")
} else {
    print("x is equal to 3")
}
```
```{r}
message <- "I change based on a condition."
if (TRUE) {
    message <- "I execute this when true!"
} else {
    message <- "I execute this when false!"
}

print(message)
```

# 7. Comparison Operators
Here is a list of some handy comparison operators and their syntax:

- Less than: <
- Greater than: >
- Less than or equal to: <=
- Greater than or equal to: >=
- Is equal to: ==
- Is NOT equal to: !=

Comparison operators compare the value on the left with the value on the right. 
```{r}
56 >= 129
56 != 129
```

# 8. Logical Operators
There are three logical operators:

1. the AND operator (&)
2. the OR operator (|)
3. the NOT operator, otherwise known as the bang operator (!)
4. 
When we use the & operator, we are checking that two things are true
When using the & operator, both conditions must evaluate to true for the entire condition to evaluate to true and execute. Otherwise, if either condition is false, the & condition will evaluate to false and the else block will execute.
If we only care about either condition being true, we can use the | operator
When using the | operator, only one of the conditions must evaluate to true for the overall statement to evaluate to true. In the code example above, if either day == 'Saturday' or day == 'Sunday' evaluates to true the if’s condition will evaluate to true and its code block will execute. If the first condition in an | statement evaluates to true, the second condition won’t even be checked.
The ! NOT operator reverses, or negates, the value of a TRUE value.
Essentially, the ! operator will either take a true value and pass back false, or it will take a false value and pass back true.

```{r}
# AND
TRUE & TRUE
TRUE & FALSE
FALSE & FALSE

# OR
TRUE | TRUE
TRUE | FALSE
FALSE | FALSE

# NOT
!TRUE
!FALSE
```
```{r}
message <- "Should I pack an umbrella?"
weather <- "cloudy"
high_chance_of_rain <- TRUE

if (weather == "cloudy" & high_chance_of_rain == TRUE) {
    message <- "Pack umbrella!"
} else {
    message <- "No need for umbrella!"
}
```

# 9. Calling a Function
```{r}
sort(c(2, 4, 10, 5, 1))
# Outputs c(1,2,4,5,10)
length(c(2, 4, 10, 5, 1))
# Outputs 5
sum(5, 15, 10)
# Outputs 30

data <- c(120, 22, 22, 31, 15, 120)
# assigning the values to the vector data
unique_vals <- unique(data)
# returns a vector with only the unique elements in that vector (removing all duplicates); calls this function and pass in the argument data; saves the results inside the variable named unique_vals
print(unique_vals)
# prints the variable to see what is inside

solution <- sqrt(49)
# gets the square root of 49 and saves it as the variable 'solution'
print(solution)
round_down <- floor(3.14) # rounds a decimal down to the next integer
round_up <- ceiling(3.14) # rounds up to the next integer
print(round_down)
print(round_up)
```

# 10. Importing Packages
```{r}
# load libraries
library(dplyr)
library(readr)
```
```{r}
# load data frame
artists <- read_csv("artists.csv")
```
```{r}
# inspect data frame
artists
head(artists)
summary(artists)
```
```{r}
artists %>%
    select(-country, -year_founded, -albums) %>%
    filter(spotify_monthly_listeners > 20000000, genre != "Hip Hop") %>%
    arrange(desc(youtube_subscribers))
```

# Introduction to Data Frames R 
# 1. What is a Data Frame 
# - a data frame is an R object that that stores tabular data in a table structure made up of rows and columns

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r}
# load data frame
songs <- read_csv("songs.csv")
```
# OUTPUT:
# ── Column specification ────────────────────────────────────────────────────────
# cols(
  # song = col_character(),
  # artist = col_character(),
  # streams = col_double(),
  # genre = col_character()
# )
```{r}
songs
```

column1,column2,column3
value1,value2,value3
value4,value5,value6

# 2. Loading and Saving CSVs
```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r}
# load data frame
artists <- read_csv("artists.csv") # reads a csv file and returns the column specifications
artists

# write_csv(df,'new_csv_file.csv') # saves data from a data frame to a csv file
```

# 3. Inspecting Data Frames
```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv("artists.csv")
```

```{r}
# inspect data frame
artists # returns the entire data frame

head(artists) # returns only the first 6 rows of the data frame

summary(artists) # gives a summary statistics of the data frame such as mean, median, minimum, and maximum for each numeric column and class and length for non-numeric columns
```

# 4. Piping
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv("artists.csv")
```

```{r}
# inspect data frame with pipe
artists %>% # this is called piping and it helps increase readability of the data frame by piping the value on its left into the first argument of the function that follows it
    head()
# so this becomes 'head(df)'
# this is mostly important for its ability to link multiple function calls together
```

# 5. Selecting Columns
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv("artists.csv")
```

```{r}
# select one column
artist_groups <- select(artists, group)
print(artist_groups)
```

```{r}
# select multiple columns
# first method
group_info <- select(artists, group, spotify_monthly_listeners, year_founded)

# second method - piping
artists %>%
    select(group, spotify_monthly_listeners, year_founded)
```

# 6. Excluding Columns
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv("artists.csv")
```

```{r}
# select all columns except one
no_albums <- artists %>%
    select(-albums) # putting a minus will remove the column
no_albums
```

```{r}
# select all columns except a set
df_cols_removed <- artists %>%
    select(-genre, -spotify_monthly_listeners, -year_founded)
df_cols_removed
```

# 7. Filtering Rows with Logic I
