---
title: "R Notebook"
author: Caroline Sanicola
date: August 26, 2023
output: 
    html_notebook:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: true
    theme: slate
    highlight: default
---

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(ggpubr)
library(ggsci)
library(ggthemes)
library(ggExtra)
library(ggcorrplot)
library(ggfortify)
library(ggrepel)
```

# 1. Calculations
```{r}
25 * 4 + 9 / 3
```

# 2. Comments
```{r}
# this is how to comment in an R code
3 * 3 * 3
```

# 3. Data Types
```{r}
class(2) # numeric
class("hello") # character
class("23") # character
class(FALSE) # logical
class(NA) # logical
```

# 4. Variables
We assign a value to a variable using the assignment operator `<-`
```{r}
x <- 3
x

full_name <- "Caroline Sanicola"
print(full_name)

# Greeting
message_string <- "Hello there"
print(message_string)

# Farewell
message_string <- "Hasta la vista"
print(message_string)
```

# 5. Vectors
List-like structure that contain items of the same data type.
```{r}
spring_months <- c("March", "April", "May", "June")
spring_months

# Vector of numbers
numbers <- c(1, 2, 3, 4, 5)
numbers
```

# 6. Conditionals
```{r}
if (TRUE) {
    print("Hello")
}

if (FALSE) {
    print("Hello")
}

# if else
if (TRUE) {
    print("Hello")
} else {
    print("Goodbye")
}

# if else if
if (TRUE) {
    print("Hello")
} else if (FALSE) {
    print("Goodbye")
} else {
    print("See you later")
}
```
```{r}
# if statement
x <- 5
if (x > 3) {
    print("x is greater than 3")
}

# if else statement
x <- 2
if (x > 3) {
    print("x is greater than 3")
} else {
    print("x is less than or equal to 3")
}

# if else if statement
x <- 3
if (x > 3) {
    print("x is greater than 3")
} else if (x < 3) {
    print("x is less than 3")
} else {
    print("x is equal to 3")
}
```
```{r}
message <- "I change based on a condition."
if (TRUE) {
    message <- "I execute this when true!"
} else {
    message <- "I execute this when false!"
}

print(message)
```

# 7. Comparison Operators
Here is a list of some handy comparison operators and their syntax:

- Less than: <
- Greater than: >
- Less than or equal to: <=
- Greater than or equal to: >=
- Is equal to: ==
- Is NOT equal to: !=

Comparison operators compare the value on the left with the value on the right. 
```{r}
56 >= 129
56 != 129
```