---
title: "R Notebook"
author: Caroline Sanicola
date: August 26, 2023
output: 
    html_notebook:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: true
    theme: slate
    highlight: default
---

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(ggpubr)
library(ggsci)
library(ggthemes)
library(ggExtra)
library(ggcorrplot)
library(ggfortify)
library(ggrepel)
```

# Introduction to R
# 1. Calculations
```{r}
25 * 4 + 9 / 3
```

# 2. Comments
```{r}
# this is how to comment in an R code
3 * 3 * 3
```

# 3. Data Types
```{r}
class(2) # numeric
class("hello") # character
class("23") # character
class(FALSE) # logical
class(NA) # logical
```

# 4. Variables
We assign a value to a variable using the assignment operator `<-`
```{r}
x <- 3
x

full_name <- "Caroline Sanicola"
print(full_name)

# Greeting
message_string <- "Hello there"
print(message_string)

# Farewell
message_string <- "Hasta la vista"
print(message_string)
```

# 5. Vectors
List-like structure that contain items of the same data type.
```{r}
spring_months <- c("March", "April", "May", "June")
spring_months

# Vector of numbers
numbers <- c(1, 2, 3, 4, 5)
numbers
```

# 6. Conditionals
```{r}
if (TRUE) {
    print("Hello")
}

if (FALSE) {
    print("Hello")
}

# if else
if (TRUE) {
    print("Hello")
} else {
    print("Goodbye")
}

# if else if
if (TRUE) {
    print("Hello")
} else if (FALSE) {
    print("Goodbye")
} else {
    print("See you later")
}
```
```{r}
# if statement
x <- 5
if (x > 3) {
    print("x is greater than 3")
}

# if else statement
x <- 2
if (x > 3) {
    print("x is greater than 3")
} else {
    print("x is less than or equal to 3")
}

# if else if statement
x <- 3
if (x > 3) {
    print("x is greater than 3")
} else if (x < 3) {
    print("x is less than 3")
} else {
    print("x is equal to 3")
}
```
```{r}
message <- "I change based on a condition."
if (TRUE) {
    message <- "I execute this when true!"
} else {
    message <- "I execute this when false!"
}

print(message)
```

# 7. Comparison Operators
Here is a list of some handy comparison operators and their syntax:

- Less than: <
- Greater than: >
- Less than or equal to: <=
- Greater than or equal to: >=
- Is equal to: ==
- Is NOT equal to: !=

Comparison operators compare the value on the left with the value on the right. 
```{r}
56 >= 129
56 != 129
```

# 8. Logical Operators
There are three logical operators:

1. the AND operator (&)
2. the OR operator (|)
3. the NOT operator, otherwise known as the bang operator (!)
4. 
When we use the & operator, we are checking that two things are true
When using the & operator, both conditions must evaluate to true for the entire condition to evaluate to true and execute. Otherwise, if either condition is false, the & condition will evaluate to false and the else block will execute.
If we only care about either condition being true, we can use the | operator
When using the | operator, only one of the conditions must evaluate to true for the overall statement to evaluate to true. In the code example above, if either day == 'Saturday' or day == 'Sunday' evaluates to true the if’s condition will evaluate to true and its code block will execute. If the first condition in an | statement evaluates to true, the second condition won’t even be checked.
The ! NOT operator reverses, or negates, the value of a TRUE value.
Essentially, the ! operator will either take a true value and pass back false, or it will take a false value and pass back true.

```{r}
# AND
TRUE & TRUE
TRUE & FALSE
FALSE & FALSE

# OR
TRUE | TRUE
TRUE | FALSE
FALSE | FALSE

# NOT
!TRUE
!FALSE
```
```{r}
message <- "Should I pack an umbrella?"
weather <- "cloudy"
high_chance_of_rain <- TRUE

if (weather == "cloudy" & high_chance_of_rain == TRUE) {
    message <- "Pack umbrella!"
} else {
    message <- "No need for umbrella!"
}
```

# 9. Calling a Function
```{r}
sort(c(2, 4, 10, 5, 1))
# Outputs c(1,2,4,5,10)
length(c(2, 4, 10, 5, 1))
# Outputs 5
sum(5, 15, 10)
# Outputs 30

data <- c(120, 22, 22, 31, 15, 120)
# assigning the values to the vector data
unique_vals <- unique(data)
# returns a vector with only the unique elements in that vector (removing all duplicates); calls this function and pass in the argument data; saves the results inside the variable named unique_vals
print(unique_vals)
# prints the variable to see what is inside

solution <- sqrt(49)
# gets the square root of 49 and saves it as the variable 'solution'
print(solution)
round_down <- floor(3.14) # rounds a decimal down to the next integer
round_up <- ceiling(3.14) # rounds up to the next integer
print(round_down)
print(round_up)
```

# 10. Importing Packages
```{r}
# load libraries
library(dplyr)
library(readr)
```
```{r}
# load data frame
artists <- read_csv("artists.csv")
```
```{r}
# inspect data frame
artists
head(artists)
summary(artists)
```
```{r}
artists %>%
    select(-country, -year_founded, -albums) %>%
    filter(spotify_monthly_listeners > 20000000, genre != "Hip Hop") %>%
    arrange(desc(youtube_subscribers))
```

# Introduction to Data Frames R 
# 1. What is a Data Frame 
# - a data frame is an R object that that stores tabular data in a table structure made up of rows and columns

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r}
# load data frame
songs <- read_csv("songs.csv")
```
# OUTPUT:
# ── Column specification ────────────────────────────────────────────────────────
# cols(
  # song = col_character(),
  # artist = col_character(),
  # streams = col_double(),
  # genre = col_character()
# )
```{r}
songs
```

column1,column2,column3
value1,value2,value3
value4,value5,value6

# 2. Loading and Saving CSVs
```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r}
# load data frame
artists <- read_csv("artists.csv") # reads a csv file and returns the column specifications
artists

# write_csv(df,'new_csv_file.csv') # saves data from a data frame to a csv file
```

# 3. Inspecting Data Frames
```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv("artists.csv")
```

```{r}
# inspect data frame
artists # returns the entire data frame

head(artists) # returns only the first 6 rows of the data frame

summary(artists) # gives a summary statistics of the data frame such as mean, median, minimum, and maximum for each numeric column and class and length for non-numeric columns
```

# 4. Piping
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv("artists.csv")
```

```{r}
# inspect data frame with pipe
artists %>% # this is called piping and it helps increase readability of the data frame by piping the value on its left into the first argument of the function that follows it
    head()
# so this becomes 'head(df)'
# this is mostly important for its ability to link multiple function calls together
```

# 5. Selecting Columns
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv("artists.csv")
```

```{r}
# select one column
artist_groups <- select(artists, group)
print(artist_groups)
```

```{r}
# select multiple columns
# first method
group_info <- select(artists, group, spotify_monthly_listeners, year_founded)

# second method - piping
artists %>%
    select(group, spotify_monthly_listeners, year_founded)
```

# 6. Excluding Columns
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv("artists.csv")
```

```{r}
# select all columns except one
no_albums <- artists %>%
    select(-albums) # putting a minus will remove the column
no_albums
```

```{r}
# select all columns except a set
df_cols_removed <- artists %>%
    select(-genre, -spotify_monthly_listeners, -year_founded)
df_cols_removed
```

# 7. Filtering Rows with Logic I
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv("artists.csv")
```

```{r}
# filter rows one condition
rock_groups <- artists %>%
    filter(genre == "Rock")

rock_groups
```

```{r}
# filter rows multiple conditions
popular_rock_groups <- artists %>%
    filter(genre == "Rock", spotify_monthly_listeners > 20000000)

popular_rock_groups
```

# 8. Filtering Rows with Logic II
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv("artists.csv")
```

```{r}
# filter rows with or
korea_or_before_2000 <- artists %>%
    filter(country == "South Korea" | year_founded < 2000) # using the OR operator

korea_or_before_2000
```

```{r}
# filter rows with not !
not_rock_groups <- artists %>%
    filter(!(genre == "Rock")) # the not operator needs to be in paranthesis before the filter argument

not_rock_groups
```

# 9. Arranging Rows
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv("artists.csv")
```

```{r}
# arrange rows in ascending order
group_asc <- artists %>%
    arrange(group) # by default the arrange will sort in ascending order

group_asc
```

```{r}
# arrange rows in descending order
youtube_desc <- artists %>%
    arrange(desc(youtube_subscribers)) # to sort by descending, it needs to be specified

youtube_desc
```

# 10. Review
---
title: "Introduction to Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
artists <- read_csv("artists.csv")
```

```{r}
# select columns
chosen_cols <- artists %>%
    select(-country, -year_founded, -albums)
head(chosen_cols)
```

```{r}
# filter rows
popular_not_hip_hop <- chosen_cols %>%
    filter(spotify_monthly_listeners > 20000000, genre != "Hip Hop")
head(popular_not_hip_hop)
```

```{r}
# arrange rows
youtube_desc <- popular_not_hip_hop %>%
    arrange(desc(youtube_subscribers))
head(youtube_desc)
```

```{r}
# select columns, filter and arrange rows
# this code does all the above in one code block
artists <- artists %>%
    select(-country, -year_founded, -albums) %>%
    filter(spotify_monthly_listeners > 20000000 & !(genre == "Hip Hop")) %>%
    arrange(desc(youtube_subscribers))
head(artists)
```

# Modifying Data Frames in R
# 1. Adding a Column
---
title: "Modifying Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
dogs <- read_csv("dogs.csv")
```

```{r}
# inspect data frame
head(dogs)
```

```{r}
# add average height column
dogs <- dogs %>%
    mutate(avg_height = (height_low_inches + height_high_inches) / 2) # the mutate function takes a name-value pair as an argument and will add a new column depending on the argument you are creating
head(dogs)
```

# 2. Adding Multiple Columns
---
title: "Modifying Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
dogs <- read_csv("dogs.csv")
```

```{r}
# inspect data frame
head(dogs)
```

```{r}
# add average height, average weight and rank change columns
dogs <- dogs %>%
    mutate(
        avg_height = (height_low_inches + height_high_inches) / 2,
        avg_weight = (weight_low_lbs + weight_high_lbs) / 2,
        rank_change_13_to_16 = rank_2016 - rank_2013
    )
head(dogs)
```

# 3. Transmute Columns
---
title: "Modifying Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
dogs <- read_csv("dogs.csv")
```

```{r}
# inspect data frame
head(dogs)
```

```{r}
# update the function call to drop all existing columns
dogs <- dogs %>%
    # transmute will add the new created columns while dropping all of the rest of the columns
    transmute(
        breed = breed, # by saying this, it will keep the column from being dropped with transmute
        avg_height = (height_low_inches + height_high_inches) / 2,
        avg_weight = (weight_low_lbs + weight_high_lbs) / 2,
        rank_change_13_to_16 = rank_2016 - rank_2013
    )
head(dogs)
```

# 4. Renaming Columns
---
title: "Modifying Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
dogs <- read_csv("dogs.csv")
```

```{r message=FALSE}
# add columns and remove existing columns
dogs <- dogs %>%
    transmute(
        breed = breed,
        avg_height = (height_low_inches + height_high_inches) / 2,
        avg_weight = (weight_low_lbs + weight_high_lbs) / 2,
        rank_change_13_to_16 = rank_2016 - rank_2013
    )
```

```{r}
# check column names
original_col_names <- colnames(dogs)
print(original_col_names)
```

```{r}
# rename data frame columns
dogs <- dogs %>%
    rename(
        avg_height_inches = avg_height,
        avg_weight_lbs = avg_weight,
        popularity_change_13_to_16 = rank_change_13_to_16
    )
head(dogs)
```

```{r}
# check column names
new_col_names <- colnames(dogs)
print(new_col_names)
```

# 5. Review
---
title: "Modifying Data Frames in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r message=FALSE}
# load data frame
dogs <- read_csv("dogs.csv")
```

```{r}
# inspect data frame
head(dogs)
```

```{r}
# add new columns, drop existing columns and arrange
dogs <- dogs %>%
    transmute(
        breed = breed,
        height_average_feet = ((height_low_inches + height_high_inches) / 2) / 12,
        popularity_change_15_to_16 = rank_2016 - rank_2015
    ) %>%
    arrange(desc(popularity_change_15_to_16))
head(dogs)
```

# Data Cleaning in R
# 1. Dealing with Multiple Files
---
title: "Data Cleaning in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(readr)
library(dplyr)
```

```{r}
# list files
student_files <- list.files(pattern = "exams_.*csv") # since all of the files start with 'exams_' we will create a pattern to look for all the files that match it
print(student_files)
```

```{r message=FALSE}
# read files
df_list <- lapply(student_files, read_csv) # reads each file and applites it to a dataframe
```

```{r}
# concatenate data frames
students <- bind_rows(df_list) # concates all the data frames into one dataframe
head(students)
```

```{r}
# number of rows in students
nrow_students <- nrow(students) # counts the number of rows in the dataframe
print(nrow_students)
```

# 2. Reshaping your Data
---
title: "Data Cleaning in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(dplyr)
library(tidyr)
```

```{r}
# load students data frame
load("students.Rda")
```

```{r}
# original column names
original_col_names <- colnames(students)
print(original_col_names)
```

```{r}
# gather columns
students <- students %>%
    gather("fractions", "probability", key = "exam", value = "score")
# students: the data frame that I want to gather, which can be piped into gather()
# 'fractions' and 'probability': the columns of the old data frame that I want to turn into variables
# key: what to call the column of the new data frame that stores the variables
# value: what to call the column of the new data frame that stores the values
head(students)
```

```{r}
# updated column names
gathered_col_names <- colnames(students)
print(gathered_col_names)
```

```{r}
# unique value counts of exam
exam_counts <- count(students, exam) # takes a data frame and a column as arguments and returns a table with counts of the unique values in the named column
print(exam_counts)
```

# 3. Dealing with Duplicates
---
title: "Data Cleaning in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(dplyr)
library(tidyr)
```

```{r}
# load students data frame
load("students.Rda")
```

```{r}
# drop id column
students <- students %>%
    select(-id)
head(students)
```

```{r}
# find duplicated rows
duplicates <- students %>% duplicated() # this function checks for any duplicates in the rows (this is for rows with all the same values as another row but you can also specify a subset by doing 'distinct(item,.keep_all=TRUE)'
```

```{r}
# count duplicated rows
duplicate_counts <- duplicates %>% # this counts the amount of rows that are duplicates
    table() # this returns the results in a table format
print(duplicate_counts)
```

```{r}
# remove duplicated rows, keep only unique rows
unique_students <- students %>%
    distinct() # function to only keep the unique rows in the data frame
head(unique_students)
```

```{r}
# find and count duplicated rows in updated data frame
updated_duplicates <- unique_students %>%
    duplicated() %>%
    table()
print(updated_duplicates)
```

# 4. Splitting By Index
---
title: "Data Cleaning in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(dplyr)
library(stringr)
```

```{r}
# load students data frame
load("students.Rda")
```

```{r}
# print columns of students
print(colnames(students))
```

```{r}
# view head of students
head(students)
```

```{r}
# add gender and age columns
students <- students %>%
    mutate(
        gender = str_sub(gender_age, 1, 1), # this is a way to extract parts of the text in the column to create a new column; the characters start at index 1 so this is saying to only take the first character
        age = str_sub(gender_age, 2)
    ) # this is saying take the characters started from index 2 until the last character
head(students)
```

```{r}
# drop gender_age column
students <- students %>%
    select(-gender_age)
head(students)
```

# 5. Splitting By Character
---
title: "Data Cleaning in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(dplyr)
library(tidyr)
```

```{r}
# load students data frame
load("students.Rda")
```

```{r}
# view the head of students
head(students)
```

```{r}
# separate the full_name column
students <- students %>%
    separate(full_name, c("first_name", "last_name"), " ", extra = "merge") # takes the values in a column and splits them into new columns depending on the character
# separate(col_to_separate, c('new_col_name_1','new_col_name_2'), 'character_to_split_on', extra ='merge')
head(students)
```

# 6. Looking at Data Types
---
title: "Data Cleaning in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(dplyr)
```

```{r}
# load students data frame
load("students.Rda")
```

```{r}
# print structure of students
str(students) # gets the data types of the table
```

```{r}
# mean of age column
students %>%
    summarise(mean_score = mean(score)) # without converting the scores to numeric values (currently string values), it will return an error that the arument is not numeric or logical
```

# 7. String Parsing
---
title: "Data Cleaning in R"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
# load libraries
library(dplyr)
```

```{r}
# load students data frame
load("students.Rda")
```

```{r}
# view head of students
head(students)
```

```{r}
# remove % from score column
students <- students %>%
    mutate(score = gsub("\\%", "", score)) # the column currently has the scores formatted as '69%' so the gsub() argument removes the characters from the column
head(students)
```

```{r}
# change score column to numeric
students <- students %>%
    mutate(score = as.numeric(score)) # transforms the column from character to numeric
head(students)
```

# Visualization with R 
# - the data or the actual information you wish to visualize
# - the geometries, shortened to "geoms", describe the shapes that represent our data. Whether it be dots on a scatter plot, bars on a bar chart, or lines on a line graph. The list goes on. Geoms are the shapes that "map" our data.
# - the aesthetics, or the visual attributes of the plot, including the scales on the axes, the color, the fill, and other attributes concerning appearance.
